Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-08-15 papers/integration/omm_integ.tex
Work in progress on paper, am now going to switch to sharelatex for the rest of the paper development

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \mycite{alex_cope_integrative_2015},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade~(\cite{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}).
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

@@

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition and a
dopamine parameter dependent modulation of the input activation, so
that their equations for $a_{in}$ are:
\[
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\]
for D1MSN, and
\[
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\]
for D2MSN, where $d$ is a dopamine parameter. Increasing dopamine from
0 to 1 enhances the activation in the D1 model neurons, whereas it
decreases the activation of the D2 model neurons, in line with
experimental observations~\cite{refs}. Note that the equation for
$a_{in}^{D1}$ differs from that used in the \ccg~model, for which the
cortico-striatal weights are multiplied by $(1+d)$ rather than $(0.2+d)$

The equations given above apply to each element in the population
whose behaviour is defined by the component. The value of the
activation $A$ (and where relevant the shunting input, $S$, also) is
determined by summing the weighted inputs to the population:
\[
A = \sum_{i}w_i^{act} x_i^{act}
\]
\[
S = \sum_{i}w_i^{sh} x_i^{sh}
\]

@@

The magnification factor, $M_f$, gives the relationship between the
radial angle $r$ and the retinal neural density. It is given by:
\[
   M_f = {W_{nfs} \over E_2 ln ({W_{fov} \over 2 E_2} + 1)}
\]
where $W_{nfs}$ is the width of the retinotopic neural fields (50),
$W_{fov}$ is the eye's field of view, which is set to 61 degrees, a
reduction from the biophysically accurate 150 degrees due to the small
number of neurons in the retinotopic neural field. $E_2$ is the
eccentricity (in degrees) at which the retinal neural density has
halved compared with the foveal region.

@@

% Brain to SBG, seeing as that's how I explained it above.
A number of studies have considered the form of the connection between
the deeper layers of the superior colliculus and the saccadic burst
generator~\cite{van_gisbergen_experimental_1985,ottes_visuomotor_1986,waitzman_superior_1991,groh_converting_2001,arai_model_2004,goossens_dynamic_2006,tabareau_geometry_2007,van_opstal_linear_2008,goossens_optimal_2012},
which has become known as the spatial temporal transform (STT).  The
spatial part of the STT is thought to be implemented by a
weight-mapping~\cite{tabareau_geometry_2007,arai_model_2004} and we
follow this idea.

% Words along the lines of:
Arai and co-workers trained their neural network model to discover the
weight map~\cite{arai_model_2004}.

% Then...
Tabareau and co-workers wrote down the theoretical form of the weight
map and found it matches well to the results of Arai et
al.~\cite{tabareau_geometry_2007}.

The approach of Arai~et.~al. was
not feasible with the present model due to the length of time required
to run the model, its stochasticity and the increased size of the maps
here compared with those used by Arai~et.~al.

>>>>

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \mycite{cope_basal_2017},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade~(\cite{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}).
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

@@

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition and a
dopamine parameter that modulates the input activation, so
that their equations for $a_{in}$ are thus:
\[
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\]
\[
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\]
where $d$ is a dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases
the activation of the D2 model elements, in line with experimental
observations~\cite{refs}. Note that the equation for $a_{in}^{D1}$
differs from that used in the \ccg~model, for which the
cortico-striatal weights are multiplied by $(1+d)$ rather than
$(0.2+d)$.

The equations given above are applied to each element in a
population. The value of the activation $A$ (and where relevant, the
shunting input, $S$) is determined by summing the weighted
inputs to the population:
\[
A = \sum_{i}w_i^{act} x_i^{act}
\]
\[
S = \sum_{i}w_i^{sh} x_i^{sh}
\]

@@

The magnification factor, $M_f$, gives the relationship between the
radial angle $r$ and the retinal neural density. It is given by:
\[ \label{eq:mf}
   M_f = {W_{nfs} \over E_2 ln ({W_{fov} \over 2 E_2} + 1)}
\]
where $W_{nfs}$ is the width of the retinotopic neural fields (50),
$W_{fov}$ is the eye's field of view, which is set to 61 degrees, a
reduction from the biophysically accurate 150 degrees due to the small
number of neurons in the retinotopic neural field. $E_2$ is the
eccentricity (in degrees) at which the retinal neural density has
halved compared with the foveal region.

@@

% Brain to SBG, seeing as that's how I explained it above.
A number of studies have considered the form of the connection between
the deeper layers of the superior colliculus and the saccadic burst
generator~\cite{van_gisbergen_experimental_1985,ottes_visuomotor_1986,waitzman_superior_1991,groh_converting_2001,arai_model_2004,goossens_dynamic_2006,tabareau_geometry_2007,van_opstal_linear_2008,goossens_optimal_2012},
which has become known as the spatial temporal transform (STT).  The
spatial transform is thought to be implemented by a
weight-mapping~\cite{tabareau_geometry_2007,arai_model_2004} and we
follow this idea.
% Words along the lines of:
Arai and co-workers trained a 20x20 neural network model of the
superior colliculs to discover the weight map under the assumption of
2D Gaussian activation profiles~\cite{arai_model_2004}.
%
The training approach of Arai~et.~al. was not feasible in this study
due to the length of time required to run our model and its
stochasticity, which means multiple runs of the model were necessary
in order to generate output statistics.
%
Tabareau and co-workers wrote down a theoretical form of the weight
map, which follows from the mapping of \cite{ottes_visuomotor_1986}
and the assumption of invariant 2D Gaussian activity profiles in
SC~\cite{tabareau_geometry_2007}. As they found it closely resembles the
results of Arai et al., and it is a simple formulation, we considered
it as the means to generate the six weight maps in our own model.
%
One barrier to the use of the weight map
in \cite{tabareau_geometry_2007} was the \ccg~model's violation of the
\emph{invariant integral hypothesis}. This states that the number of spikes
emitted by a neural element during a saccade (or in our model, the
integral of the neuron's output during the saccade) should be a
function only of its position within the hill of collicular
activity. That is, for any time-dependent hill of activity
$\mathcal{A}(\mathbf{z},t)$ at $\mathbf{z} = (r,\phi)$ on the collicular
surface, the integrated activity  $A_{\mathbf{x}}$ in an element at a
vector $\mathbf{x}$ away from $\mathbf{z}$ is
\[
A_{\mathbf{x}} = \int_t \mathcal{A}(\mathbf{z}-\mathbf{x}, t) dt
\]
which is invariant for all $\mathbf{z}$. However, the very mapping on
which the \cite{tabareau_geometry_2007} result is based leads to a
very \emph{variant} activity profile in the \ccg~model. A luminance of
a given size which excites activity near to the fovea causes activity
in a large number of neurons, whereas activity far from the fovea
excites a much smaller region. This effect is clearly demonstrated in
Fig.~\ref{fig:mapping} for equal sized targets both on and distal from
the fovea.

This led us to hypothesize that the retinotopic mapping be accompanied
by an associated widening projection field such that the hill of
activity in superior colliculus is invariant with position on the
collicular surface. There are a number of locations in the system in
which this widening projection field could exist. It could be
implemented in the projections between the retinal populations and the
superficial layer of SC along with the projection between the World
and the FEF population. However, this would affect activity within the
basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010}
and \cite{vokoun_response_2014}. Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
(\ccg has two layers; SC\_sup) and SC\_deep) named SC\_deep2. We
introduced a widening projection based on a Gaussian projection field
whose width varies in inverse proportion to the magnification factor.

A further problem - the temporal profile of the hill? No - 6 channels
not 4!
